Clouds and Sensitivities Across a Hierarchy of GFDL CMIP6 Models

### 2020-11-03 Silvers_WalkerCell_response2reviewers.tex

\begin{enumerate}
  \item Section 3 is meant to focus on the impact of LWCRE, but subsequent sections frequently reference the comparison between LWCRE-on and LWCRE-off as well (i.e. lines 361-373, 398-401, 459-462).
  \textcolor{blue}{I have added this sentence to the end of the introduction where the layout of the paper is described: 'The longwave 
  cloud radiative effect (LWCRE) is an important element for all of the experiments and will be discussed throughout the paper.'  My 
  response to your next point is also relevant to this point.   During the writing I attempted consolidating the comparison between 
  LWCRE on/off to one section but found it to 
  create more problems than it solved because the LWCRE on/off experiments have such a strong influence on all of the other   
  sections.  You mention lines 361-373, which happens to be a perfect example of the difficulty.  Figures 5 and 6 illustrate the 
  overlapping influence of the LWCRE on/off effects and domain size dependence.  As a result, I have chosen to largely focus on your 
  other excellent recommendations for reworking the organization of the manuscript.}
  \item Comparisons between E25 and P25 are made in multiple sections, and some of these results are referenced in passing (i.e. lines 518-519, "It is also worth noting that in contrast to the P25 case which has strong domain shear..."). Given the importance of this resolution in the manuscript to link GCMs and CRMs, I believe that this deserves its own section.
  \textcolor{blue}{Your review has convinced me.  I have reorganized the paper to include a new subsection 3.2 in which P100, P25, and E25 are compared to highlight the impact that resolution and the convective parameterization have.  I think this makes the overall 
  comparison between GCM-like and CRM-like models more logical.   The new subsection shows domain mean profiles of RH, domain mean condensate, cloud fraction, atmospheric heating, and temperature.  I have also included a figure with additional Hovmoller precipitation diagrams focusing on P25 and E25. To further aid in the comparison between E25 and P25, the P25 control and P25 LWCRE-off profiles of total condensate have been added to Figure 11 (note the number of this figure will change depending on the E25/P25 section).  An additional panel was added to that figure showing the vertical distribution of specific humidity.  The color used to show the E25 profiles has also been changed from yellow to purple in several of the figures and I think this also helps readers to compare E25 and P25.}
  \item Figure 10, in the context of its section on resolution dependence, seems to convey a similar picture as Figure 9, and its already been well-established that removing LWCRE seems to broaden/weaken the mass circulation robustly.

@@

  \textcolor{blue}{You are right that there is some redundancy of message in Figure 10, but I think there is enough interest in the results of the 1 and 2km simulations to warrant inclusion of Figure 10. }

  \item I believe that Sections 6 and 7 can be consolidated and combined into one, given that the first half of Section 6 is already a well-constructed and well-divided summary of results.
  \textcolor{blue}{Section 6 and 7 have been combined and reorganized into one section with two subsections, one with results and one with discussion.  I think this improves the clarity.  Please see the revised manuscript.}
  \item In several parts of the manuscript, multiple references are made to Figures/Tables that appear 10 or more pages earlier. For ease of reading, perhaps quantities such as the domain-mean precipitation can be incorporated into the Figures. For example, Figure 2's legend can list the domain-mean precipitation values next to the labels.
  \textcolor{blue}{I have added domain-mean values to the precipitation shown in Figure 3, and extended the markers at the end of the time series to make Figure 3 easier to read.  
  In response to several comments from you and the other reviewers, section 3, 6, and 7 have been somewhat reorganized and hopefully make the manuscript easier to read. }
\end{enumerate}

@@

\begin{itemize}
  \item Given the overarching goal of linking CRMs with GCMs, it is understandable that a 25 km resolution is chosen for the sensitivity experiment of using parameterized vs. explicit convection, instead of a finer spacing. I would also be interested to see this comparison at resolutions representative of operational NWP models, roughly 10 km or so. There are several key differences noted in the manuscript between P25 and E25, so performing a similar comparison would be useful to assess the robustness of the results, while also providing potential relevance for regional climate modeling efforts. This is more out of my own curiosity, rather than something I request be done within the scope of this manuscript.

  \textcolor{blue}{This is a good point, and I think that we have similar interests in some of what the paper has not covered.  As you could probably see in the paper, organizing the results was challenging, and interpretation was made difficult by many unexpected complications.  Based on the number of experiments which I was already trying to organize in the writing of the manuscript I chose to keep the focus of the paper rather small.   Another factor which limited the number of experiments that I could include in the paper is that I also performed a number of uninteresting or unsuccessful experiments in which various parameter settings were modified, and in which some minor changes were made in the code of the Tiedtke large scale cloud scheme.  I have more experiments than I can handle! }

  \item Have any additional simulations been conducted at resolutions between 2-25, and 25-100 km? It would be interesting to see how smooth the transition is between these two regimes of cloud distribution and radiative response, and if a specific threshold could be found, holding all other conditions equal, to significantly affect the cloud distribution.

  \textcolor{blue}{Because this study focuses on LWCRE, convective parameterizations, and domain size in addition to grid-spacing I have not performed experiments at other resolutions (a few at 5km, at the very beginning with different code base, but the analysis was not carried very far).  Also see answer to you comment just before this one. }

  \item Some figures are a bit difficult to read, in particular Figures 6 and 13. In particular, for Figure 6 it is difficult to compare the P25 and P25L simulations because the colors on the lines are the same. Figure 13 would be best oriented as a 2X2 window similar to Figure 12, especially given the subtle differences in the rightmost panel.
  \textcolor{blue}{For most of the figures (3, 6, 7, 11,12, and 13) the labels and axis font size has been increased.  
  Figure 6 has now been split into two figures, one showing results on the small/control domain size and the other showing 
  results on the large domain.  Figure 13 have been shift to a 2x2 panel format, the color used for E25 is now purple, and the 
  title are clearer.}

  \item For each of the figures, especially those later in the manuscript, it would be helpful to make the font size on the axes a bit larger.
  \textcolor{blue}{The font size has been increased.}

@@

  \item Line 217-220, 276, 302, 485-494: While initially denoting the simulations as "CRM-like" and "GCM-like", there are some inconsistencies later in the manuscript where these are referred to as just "GCMs and CRMs". I would suggest sticking with the latter naming scheme for simplicity, making sure to state the caveats in Section 2 as you already did for the CRM-like cases.

  \textcolor{blue}{I have gone back and forth about what terminology to use.  I hesitate to call a doubly periodic simulation a GCM but you can see the difficulties that arise with the terminology that appends ‘-like’ to the GCMs and CRMs.  For now I have added ‘-like’ in the places that you noticed inconsistencies. }

  \item Line 321-323: "...but also due to an interaction between the convective parameterization and the LWCRE." If this is known, can you elaborate on it? I take this to be the cause of the domain-mean shear alluded to in lines 315-316, and believe this explanation would strengthen the comparison.

  \textcolor{blue}{The new subsection, 3.2 should clarify this, especially Figure 5. In contrast to the P25 (lwcre-on) simulation (Figure 1, left; Figure 4d), the initial several months of both P25 
  (lwcre-off; Figure 5d) and E25 (lwcre-on; Figure 8b) show a precipitation maximum that is centered closely on the SST 
  maximum in the middle of the domain.  The steady state solutions from the equilibrated period of these simulations (Figure 1) are 
  also both more symmetric about the center of the domain.  This implies that the source of asymmetry could be either the lwcre, 
  or the parameterized convection.  However, both P25(lwcre-off) and E25(lwcre-on) eventually transition to a much more irregular evolution of precipitation (see Figure 5d after 600 days) in which the precipitation maximum can reside well off of the 
  precipitation maximum for several months at a time.  However, both of these simulations have a more symmetric time-averaged distribution than the P25 case.  I then made the inference that it is the interaction between the two that leads to an asymmetric distribution of the steady state. }

\textcolor{blue}{
This has been concluded mostly because the simulation is markedly more symmetrical when the LWCRE is off but the convection parameterization is still on, and the equilibrium state is more symmetrical when the LWCRE is on but the convection parameterization is off (Figure 1).  The time evolution of precipitation for the LWCRE-off, Conv-off (E25 lwcre off) also shows regular oscillations (large) about the SST maximum in a way that is distinct from either P25 lwcre-off or E25 lwcre-on. }

  \item Line 326-343: The focus in Table 2 is on the equilibrium state. Is a time evolution of subsidence fraction available? I'd be curious to see how SF changes from the early stages of the simulation, i.e. if it's becoming more or less aggregated over time. The somewhat inconsistent comparison between LWCRE-on and LWCRE-off experiments here is noteworthy, given that LWCRE is arguably the most important mechanism for maintenance of the aggregated state in non-rotating RCE studies.
\textcolor{blue}{Good question, please see the included Figure 1.  Once the experiments equilibrate the subsidence fraction (SF) is quite steady in time, despite rather large variability (on the order of 0.1-0.15).   Variability increases with grid-spacing to the point that for P100 experiments averages must be taken over at least 3-5 months to obtain the long term mean SF.
Figure \ref{fig:subsfracTS}  
shows the subsidence fraction for the control (LWCRE-on) experiments.  The LWCRE-off cases are also very steady in time with a similar degree of variability.  The mean 
differences of SF between the LWCRE on/off experiments are shown in an updated Table 2.  It should also be noted that SF in the case of the Walker 
Circulation should be expected to be relatively large, and more consistent across experiments (relative to pure RCE) because of the prescribed SST warm 
patch that will largely fix the location of deep convection to the middle of the domain and force an overturning circulation.  
The values used in Table 2 for the originally submitted manuscript were computed 
using the time averaged mid-tropospheric vertical velocity because I had neglected to save vertical velocity at a daily frequency.  However, I have since 
found that the daily vertical pressure velocity (omega) was saved on the 500 hPa pressure level.  I have recomputed the SF values using these omega values
and updated Table 2 (caption has been updated).  The SF values now consistently show that the LWCRE-off experiments have a lower SF, the values are more consistent across experiments, and the surprisingly large difference between the LWCRE off/on experiments for the P100 large domain experiments is reduced to a value of 0.1.  I have deleted 'with the exception of E25' from Line 332 of the submitted manuscript.}

@@

\begin{figure}
  \centering
    \includegraphics[width=0.8\columnwidth]{walkerfigs/SubsFrac_TS.eps}  
  \caption{Subsidence Fraction over the first two years of simulation (E1 and E2 only extend for 6 months.).  Values computed from the pressure 
  velocity on the 500 mb pressure surface.}
  \label{fig:subsfracTS}
\end{figure}

  \item Line 374-378: "An additional unexpected change that results from increasing the domain size is an upward shift of the cloud fields..." This is an interesting result! Do you have any speculation as to why this happened?
  \textcolor{blue}{When I first saw this upward shift I thought it must be an error in some part of the setup or analysis... when I convinced myself it was not an error I was fascinated.  However, it is one of several interesting results that I have resisted
  pursuing in this paper for fear of being pulled in two many directions.  My speculation is that this is due to a difference of mixing between the BL and the lower troposphere for the large domain experiments, there are much higher values of specific humidity in the BL for the P100L.  Perhaps with a larger physical extent of the SST warm patch there is simply more convective mixing?}

@@

Accurate Key Points: Yes

The present manuscript investigates the effect of longwave cloud radiative effect and spatial resolution on the mock-
Walker circulation in a model that can consistently switch between explicit and parametrized convection. More specifically,
the authors design a series of seven experiments using the Geophysical Fluid Dynamics Laboratory (GFDL) general circu-
lation model (GCM) AM 4.0. with prescribed surface temperature and different domain sizes, grid spacing, active/inactive
longwave cloud radiative effect (LWCRE), and explicit/parametrized convection (Section 2). They find that LWCRE orga-
nizes the Walker cell’s mean precipitation and circulation (Section 3), and increases large-scale precipitation while domain
size leads to a more modest increase in precipitation as the condensate peak shifts from high-level clouds to low-level clouds
(Section 4). In contrast, explicit convection tends to lead to more high-level clouds and more convective aggregation, i.e.
stronger circulation, thermodynamic spatial gradients, smaller precipitation mean but larger precipitation variance, and
lower surface enthalpy fluxes, a lot of it forced by LWCRE (Section 5). The findings are then thoroughly summarized in
the discussion (Section 6) and conclusion (Section 7) sections.

This series of mock-Walker cell simulation is new, the manuscript is well-written, the results are consistent with past
literature, and the topic is of interest to the climate community as it fits well within the “climate model hierarchy” [3, 1],
all of which warrant eventual publication in the Journal of Advances in Modeling Earth Systems. I list a few comments
below and would be happy to help with another round of minor revisions if necessary.

\textcolor{blue}{Thank you for your review and the encouraging synopsis, it has helped me to improve the paper.  I think that 
the changes made as a result of the comments from you and the other reviewers has substantially improved the clarity of the 
results and descriptions.  I have incorporated many of your suggestions, in particular a panel looking at the vertically resolved 
specific humidity and a simple look at observations from MISR.}

@@

\subsection{Major comments}
\subsubsection{Why limit the investigation of cloud radiative effect to LWCRE?}
While the climate dynamics and self-aggregation literature consistently find that the LWCRE is the most important
radiative forcing when it comes to convective aggregation (e.g. [6, 4, 2]), the consensus is less clear when it comes to
shortwave cloud radiative effect (SWCRE). Given the clean simulation setup and the insight obtained from the LWCRE-off
simulations, \textbf{would it be possible to repeat the experiments with SWCRE and/or the full cloud radiative
effect (CRE) turned off?}

I would recommend to:

\begin{itemize}
  \item Rerun at least one simulation with SWCRE-off or the full CRE-off, or
  \item Thoroughly justify why the authors decided to focus on LWCRE only rather than the full CRE. The paragraph
currently introducing the LWCRE-off experiments (L202) does not justify why the experiment is done in the first
place, even if the manuscript’s results confirm the importance of this experiment a posteriori.
\textcolor{blue}{Good question.  It is possible with this configuration to repeat experiments in which the SWCRE is turned off/on.  
We are using a very similar physics code as that used by \textit{Popp \& Silvers, 2017, J. of Clim.}.   However, we have chosen to limit ourselves in this case to the 
LWCRE experiments for several reasons, the foremost being simplicity of focus.  In addition, the current study varies LWCRE, resolution, domain size, and 
the convective parameterization.  As a result, organizing the current range of results is already a challenge and we chose not to 
expand the study further.  I have no doubt that the impact of the SWCRE on the coupling between the LS circulation and clouds will lead 
to more insight and I hope that it will be investigated in future studies.  Additional reasons for focusing our study on the LWCRE on/off experiments
are as follow.  First, this provides a clean overlap with CFMIP experiments which emphasize LWCRE over the SWCRE due to the complicating impacts of
SWCRE on the surface energy budget.  Second, there are numerous studies which have shown that the atmospheric shortwave heating is 
much smaller than the longwave heating (e.g. Pendergrass \& Hartmann, 2014; Popp \& Silvers, 2017 (their Figures 1 \& 4)).}

\textcolor{blue}{The following has been added to the text: One of the ways in which clouds influence the flow of energy through the 
atmosphere is by absorbing and 
reemitting longwave radiation....  and later in the same paragraph..... Clouds also interact with shortwave radiation.  We have chosen 
to only focus on the LWCRE for several reasons.  
On timescales of years and over large domains, the LWCRE has been shown (e.g. Popp \& Silvers, 2017) to have a much larger 
influence on the atmospheric 
heating and the large-scale circulations compared to the SWCRE.  The SWCRE is important on diurnal timescales and for the surface
energy budget.  However, we choose to leave simulations exploring the SWCRE in the context of the mock-Walker circulation
for a later study.}
\end{itemize}

@@

\subsubsection{Descriptive nature of the manuscript}

I appreciated the explicit questions from line 80 to 84, which respectively correspond to Sections 3, 4, and 5:

\begin{itemize}
  \item How do clouds influence the overturning circulation?
  \item To what extent are the deep convective clouds and the low-level clouds coupled through the overturning circulation?
  \item When simulating tropical overturning circulations, how well does a GCM compare to a cloud-resolving model (CRM)?
\end{itemize}

However, I found the manuscript to be overly descriptive at times, possibly restricting the insight readers could get from
the AM4.0 Walker setup. For each of the three questions, it could be helpful to use the simulation data to offer a simple
rationale if not a quantitative explanation for why the authors find the results summarized in Section 6.

\textcolor{blue}{Sections 6 and 7 have been combined and edited to help clarify the end of the paper.  The goal has been to 
maximize the insight that readers can get without overloading them with too many details.  I think the revised manuscript 
succeeds in this and is also more quantitative than the original submission.}

\begin{enumerate}
  \item (Section 3)
    \begin{itemize}
      \item Is it possible to simply relate the difference in circulation/vertical velocity to the difference in CRE (e.g. using the
two box model for a given value of evaporation minus precipitation)?
      \item How does CRE compare to circulation strength throughout each simulation once the simulation has equilibrium
(this would give more data points than a single mean circulation/mean CRE comparison per simulation)?
  \textcolor{blue}{These are good questions.  There is great appeal to me in the simplicity and physical reasoning that drives many of the two box models, and other idealized
  theoretical studies of overturning circulations.  However, the focus of this work has always been more on the comparison between the
  GCM-like and CRM-like configurations.  I think a comparison between a two-box model and some of these simulations would be 
  very interesting, but in my opinion I needed to get a sense of what was happening with these experiments before bringing in an 
  additional framework for comparison.  In addition, the interesting differences in the structure of precipitation, the strong influence
  on the LWCRE and the significant differences in the vertical structure of the condensate between the GCM-like and CRM-like 
  models essentially distracted me from two-box model type work that helped to motivate the original study.  It is also not immediately clear to me which two-box model, which measure of CRE and which measure of circulation strength you have in mind.  I would hope
  that multiple two-box models, various measures of CRE (TOA CRE, ACRE, or differences of particular variables between a control and LWCRE-off experiment), and various measures of the circulation strength would all point to similar insights into the underlying physics, but the ambiguity of the current questions highlights the need for a future study rather than quick inclusion in this paper.}  
  \item (Section 4, Lines 407-409)
    \begin{itemize}
      \item Could the authors show correlations or scatter plots across simulations and across time within a given simulation
to back up their claim about the link between low-level cloud fraction, LWCRE, and convective versus large-scale
fraction of total precipitation?
      \item The causality claim also seems a bit bold given that LWCRE is the only effect that the authors turned on/off; it could
be for instance that surface latent heat fluxes plays as big of a role, which could be tested by turning the dependence
of latent heat fluxes on surface wind speed on/off, or by making the surface latent heat fluxes horizontally uniform

### 2020-11-03 Silvers_WalkerCell_response2reviewers.tex

at each timestep. If the authors decide not to run additional experiments, I recommend nuancing that statement
given that Figure 7 shows that LWCRE is only responsible for part of this partitioning, and given that Figure 6 is
so hard to read (which could be easily improve by making the P25 and P25L different colors, etc.).
  \textcolor{blue}{These are good points you are making.  I have rewritten the lines in question here to be more consistent
  with what is clearly shown in this paper, and to not so strongly implicate a connection with low-level clouds.  
  I have concluded that more research is needed before drawing firm conclusions about the connection 
  between low-level clouds and the fraction of precipitation that is due to either the convection or the large-scale schemes.  In response 
  to your questions I have computed correlations in time (on monthly data because I don't have daily) between the convective precipitation and the 
  low-level clouds, as well as with the high-level clouds.  There are a lot of high correlation numbers, but clear conclusions are not obvious and I 
  would prefer to not commit to conclusions in print yet.  In short, several of the experiments (but not all) show high correlations between low-level 
  clouds and convective precip.  But the correlations with high-level clouds are also high.  In addition, the correlation is still high for the experiments 
  with the LWCRE-off.  So I am left desiring more information or experiments, but not for this paper.
  Figure 6 has been split into two panels and is now easier to read. }
  \item (Section 5)
    \begin{itemize}
      \item The logical flow of the CRM/GCM comparison is relatively easy to follow, but I was wondering if the authors
could expand on their baseline result that the CRM configuration leads to more upper-level clouds while the GCM
configuration leads to more lower-level clouds. Would it be possible to compare the vertically-resolved water species
budget across simulations to understand this difference?
      \item It is reassuring that this result is consistent with [5]; would it be possible to similarly trace back this difference to a
discrepancy in shallow cloud mixing, e.g. by comparing the vertical transport of water vapor between the boundary
layer and the lower troposphere across experiments?
    \textcolor{blue}{I have worked to clarify what was Figure 11 in the submitted manuscript.  An additional panel has been added showing 
    the vertical profiles of specific humidity, similar to figure 2 in Pauluis and Garner, 2006.  I have also noted in the caption, that the liquid
    condensate is less than 0.005 g/kg above 600 hPa.  I hesitate to add the sperate profiles of liquid and ice because when combined with 
    lwcre-off profiles, it makes for a very busy figure and I am not sure that it adds much.  }
\end{enumerate}

@@

\end{enumerate}

\subsubsection{Assessing the realism of the simulated mock-Walker cell}

As the authors point out (L67), the Walker cell setup can be compared to observations from the Pacific region. As their
results show large differences between the E25/P25 and the E2/E1 simulations (e.g. L420-421), how can the reader
assess which simulation best captures the Walker cell’s hydrological cycle?
Although a direct comparison is impossible because of the idealized mock-Walker cell setup, I would highly suggest
adding observational and/or meteorological reanalysis data to at least some of Section 5’s figures (Figure 8-13) to comment
on which configuration leads to the most observationally consistent Walker cell. To facilitate the comparison, the authors
could e.g. choose a period of time where the Pacific’s zonal surface temperature gradient is in the range of the experiments
listed in Table 1 $(\Delta SST/ \Delta x = 10^{-4} - 10^{-3} K/km).  $

\textcolor{blue}{As you imply, it is challenging to compare a mock-Walker circulation to observations.  I have added 
figure 16 which uses cloud fraction data from the MISR instrument in the tropical Pacific.  The following text was added near the 
end of section 5: }

\textcolor{blue}{In order to make a qualitative connection to observed clouds and hopefully motivate a more detailed comparative 
study we look 
at the clouds observed with the Multiangle Imaging Spectro-Radiometer (MISR) instrument on board NASA's Terra satellite (see 
Marchand et al., 2010 for details).  We use the domain (10N-35S) which was identified by Schwendike et al., 2014 as defining the 
regional Walker circulation in the Pacific.  The left panel of Figure 16 shows the cloud fraction averaged from March 2000-
November 2019 of the full domain of the Pacific regional Walker circulation.  This region includes much of the so-called warm pool 
beneath the ascending branch of the Walker Circulation as well as the eastern Pacific subsidence regions that encompass the 
stratocumulus cloud decks that are common off the west coast of South America.   To calculate profiles of the cloud fraction as a 
function of height we sum the MISR data over all optical depths  and plot the domain mean (dashed line), a region of subsidence 
($10^\circ$ N-$35^\circ$S; $120^\circ$W-$60^\circ$W, solid line), and a region that is characterized by deep 
convection ($10^\circ$N-$35^\circ$S; $120^\circ$E-$180^\circ$, dash-dot line).  Overall, both the upper- and low-level cloud fraction 
from MISR are closer to the cloud profiles simulated by the GCM-like models.  However, it should be kept in mind that there are 
several substantial differences between our mock Walker circulation and the observed Walker circulation.}

@@

\subsection{Minor comments}

You will find 15 minor comments attached to the annotated PDF of the manuscript. 
\textcolor{blue}{Thank you.  I have responded to your comments in the annotated PDF.}

Additional comments that apply
throughout the paper are listed below:
\begin{itemize}
  \item Labeling all figure panels using the conventional notation a/b/c/d (e.g. Figure 1, 2, etc.)
  \textcolor{blue}{Figures 1-15 now have panels labelled with a/b etc., when multiple panels are present.}
